Large language model (LLM) inference is dominated by repeated dense linear projections whose computational and memory demands grow sharply with model scale. This work studies whether Bit-Sliced Indexing (BSI) can serve as a practical compressed representation for LLM inference on GPUs while preserving a native bit-sliced execution path instead of reverting to dense low-precision matrix multiplication. The implementation grows out of an earlier CPU-side bitmap and Bit-Sliced Vector foundation in which hybrid bitmaps were used to construct BSI representations and to implement dot product, summation, and multiplication directly in bit-sliced form. Those early experiments showed that BSI could reduce memory usage and preserve useful algebraic structure, but they also showed that CPU execution was not competitive for the dot-product path that dominates transformer inference. That result motivated the transition to CUDA, where the same BSI semantics were retained while the execution path was redesigned around static BSI weight construction, runtime BSI query construction, and fused same-bit tensor-core kernels. The resulting system provides a complete experimental stack: CPU and CUDA BSI builders, a PyTorch replacement for \texttt{torch.nn.Linear}, kernel-only microbenchmarks, end-to-end next-token evaluation on LAMBADA, and comparisons against dense FP16 baselines. The stable CUDA baseline centers on an SM90/H100 operating point using \texttt{TM=32}, \texttt{R\_sweep=4}, chunk scaling, and fixed query/key slice counts of \((S_a,S_b)=(7,6)\). Under the principal evaluation setting (\texttt{decimal\_places=2}, \texttt{compress\_threshold=0.5}, \texttt{num\_samples=1000}, \texttt{max\_seq\_len=512}, and \texttt{fp16} base dtype), the \texttt{facebook/opt-6.7b} model decreases from \texttt{12700.031 MB} in dense form to \texttt{5022.281 MB} in BSI form under \texttt{scope=all}, while achieving \texttt{top1=0.741}, \texttt{top5=0.896}, and \texttt{66.452 ms} per-sample dot latency compared with the dense baseline at \texttt{top1=0.798}, \texttt{top5=0.929}, and \texttt{15.493 ms}. The main conclusion is that BSI representation alone is insufficient for competitive GPU inference. A generic bit-sliced storage layout still incurs substantial orchestration cost, and the remaining performance gap to dense Torch is dominated by the interaction between representation, staging, and kernel execution. The contribution of the work is therefore not a claim that BSI already outperforms dense inference, but a complete BSI-native inference stack and a defensible systems result: BSI becomes materially more competitive only when the data layout, chunk handling, and dot kernel are designed together for the target GPU architecture.

\noindent\textbf{Keywords:} bit-sliced indexing, bit-sliced vectors, large language model inference, CUDA, tensor cores, transformer compression, OPT, weighted popcount
