Large language model (LLM) inference is dominated by repeated dense linear projections whose computational and memory demands increase sharply with model scale. This work investigates whether Bit-Sliced Indexing (BSI) can serve as a practical compressed representation for LLM inference on GPUs while retaining a native bitwise execution path rather than falling back to dense low-precision matrix multiplication. The implementation targets PyTorch-based end-to-end inference for OPT-family models and uses a C++/CUDA extension that stores linear weights in BSI form, builds bit-sliced query activations on the fly, and computes matrix products using fused CUDA kernels based on binary tensor-core operations and weighted popcount accumulation. The central research question is not whether lower precision is generally faster, but whether a BSI-native execution pipeline can be made substantially more competitive through representation and kernel co-design.

The implementation establishes a complete experimental stack: CPU and CUDA BSI builders, a PyTorch replacement for \texttt{torch.nn.Linear}, kernel-only microbenchmarks, end-to-end next-token evaluation on LAMBADA, and comparisons against both dense FP16 baselines and SmoothQuant benchmark scripts. The stable fixed-bit CUDA baseline centers on an SM90/H100 path using \texttt{TM=32}, \texttt{R\_sweep=4}, chunk scaling, and a fixed \((S_a,S_b)=(7,6)\) query/key slice configuration. That baseline achieves meaningful model compression and maintains usable accuracy for moderate model sizes, but still trails dense Torch inference by a wide margin. Additional layout and staging studies were conducted to understand the design space, and several of those ideas were discarded after profiling or correctness regressions. Those negative results are still valuable because they help identify which parts of the pipeline are genuinely limiting performance.

The principal finding of the work is that BSI representation alone is insufficient for competitive GPU inference. A generic BSI storage layout, even when combined with tensor-core instructions, still incurs significant orchestration overhead and poor scheduler efficiency. Profiling of the same-bit tensor-core path showed that the implementation was not primarily DRAM-bandwidth limited; instead, it was constrained by low eligible warp count, high no-eligible stall time, and excessive shared/global memory transactions. This result materially reframed the optimization problem: the main challenge is not only bit count, but co-designing the BSI layout and the kernel’s access pattern so the representation matches the hardware execution model.

Empirically, the work demonstrates a clear compression and performance tradeoff. Under the primary evaluation setting (\texttt{decimal\_places=2}, \texttt{compress\_threshold=0.5}, \texttt{num\_samples=1000}, \texttt{max\_seq\_len=512}, \texttt{fp16} base dtype), the \texttt{facebook/opt-6.7b} model achieved a dense static size of \texttt{12700.031 MB} and a BSI static size of \texttt{5022.281 MB} under \texttt{scope=all}. In the same setting, BSI achieved \texttt{top1=0.741}, \texttt{top5=0.896}, and \texttt{66.452 ms} dot latency per sample, compared with the dense baseline at \texttt{top1=0.798}, \texttt{top5=0.929}, and \texttt{15.493 ms}. These results show that BSI provides real compression and a functioning inference path, but also that the remaining performance gap is dominated by the interaction between representation and kernel design. The defensible research contribution is therefore the identification and partial resolution of the real bottleneck: BSI-native inference becomes materially more competitive only when the data layout, chunk staging, and dot kernel are co-designed for the target GPU architecture.
