\chapter{Background and Foundations of Bit-Sliced Vector Algebra}

\section{Bitmaps, BSI, and the Origin of the Representation}
The conceptual foundation of the present implementation is a bitmap-oriented representation rather than a conventional dense tensor representation. In the bitmap view, a numeric vector is decomposed into several bitplanes or slices, and each slice is represented as a bitmap over vector positions. This idea is rooted in Bit-Sliced Indexing, where numeric attributes are encoded so that selection and aggregation can be carried out through bitwise logic. Earlier database work on hybrid bit-vectors showed that the efficiency of a bitmap representation depends not only on the number of bits stored but also on the density distribution of those bits and on the storage form selected for each slice \cite{hybridbitmap}.

The code foundation used in this project adopts that idea in the form of Bit-Sliced Vectors. Instead of storing a vector as a dense array of scalar values, the absolute-value bits are distributed across multiple bitmaps, while sign and existence information are tracked separately. Once that representation is available, arithmetic can be written in terms of bitwise intersections, unions, exclusive-or, and weighted bit counts. This algebraic viewpoint is the basis from which the later CUDA implementation was derived.

\section{CPU-Side BSI Construction in the Bitmap Library}
The CPU-side implementation in the \texttt{bsiCPP} codebase provides the semantic definition of a BSI object. Four structures are central to that implementation, and together they define the meaning that the later CUDA implementation must preserve.

\subsection{HybridBitmap}
\texttt{HybridBitmap} represents a single bitmap slice. Each slice can be stored either in compressed form or in verbatim form. The object keeps the bitmap words, a density estimate, and a flag indicating whether the stored form is verbatim. This choice is important because slice storage cost depends strongly on bit density. Sparse slices may benefit from compression, while dense or highly irregular slices may be better stored verbatim.

\subsection{BsiVector}
\texttt{BsiVector} is the abstract container for a full bit-sliced numeric vector. It maintains the vector of slice bitmaps, the number of slices, the row count, the decimal scaling metadata, and the sign and existence bitmaps. The existence bitmap records whether a value is present at a position. The sign bitmap records whether the value is negative. Together, these structures allow the representation to support signed, unsigned, and partially populated vectors.

\subsection{BsiSigned and BsiUnsigned}
The CPU implementation separates signed and unsigned construction through \texttt{BsiSigned} and \texttt{BsiUnsigned}. Nonnegative values can be represented without a negative sign contribution, while signed values require either sign-magnitude logic or two's-complement handling. In practice, transformer weights and activations require the signed path because the tensors contain both positive and negative values.

\subsection{Building Slices with \texttt{buildBsiVector} and \texttt{bringTheBits}}
The builder routines scale floating-point values by a decimal factor, round them to integers, determine how many slices are needed from the largest absolute value, and then transpose the resulting integer vector into slice-major bitmaps. The routine that performs the transpose is \texttt{bringTheBits}. For each element and for each bit position of interest, it tests whether that bit is set and writes the result into the corresponding bitmap slice. In other words, \texttt{bringTheBits} converts a list of integers into a stack of aligned bitplanes.

This construction step has two important consequences. First, it exposes a direct tradeoff between precision and the number of slices stored or processed. Second, it makes the storage cost of each slice observable. The builder counts the number of set bits per slice, estimates the slice density, and uses that information to decide whether the slice should remain verbatim or be stored in compressed form.

\section{Compression Heuristic and Existence Tracking}
Each slice is assigned a density estimate based on the fraction of one bits in that slice. The builder then applies a compression heuristic that compares the expected benefit of compressed storage against a user-provided threshold. This decision happens per slice rather than per vector. As a result, two BSI vectors with the same number of slices can occupy different amounts of memory if their slice densities differ.

The existence bitmap is equally important, even when the application seems dense. In a database-style workload it marks whether values are present or absent at each position. In the transformer setting used here, the existence bitmap is usually all ones because every weight element or activation element exists. Even then, the existence bitmap remains a semantic part of the representation and must be preserved by the builder logic.

\section{Bit-Sliced Vector Algebra from the Earlier CPU Study}
The earlier technical work on bit-sliced vector algebra extended the representation beyond storage and demonstrated several operations directly on the BSI form.

\subsection{Dot Product}
The bit-sliced dot product follows a simple but important idea. Given two bit-sliced vectors, corresponding magnitude slices are intersected through bitwise AND operations. The resulting set-bit counts determine how many vector positions contribute at a particular significance level. Those counts are then multiplied by the positional weight implied by the slice indices and accumulated into the final scalar result. Sign bitmaps determine whether a shared overlap contributes positively or negatively.

This formulation is central to the later CUDA design. The GPU kernel is not performing a dense multiply-add over scalar lanes. It is computing overlaps between bitplanes and then converting those overlaps into weighted scalar contributions.

\subsection{Vector Summation}
Bit-sliced summation uses the same principle without the second input vector. The number of set bits in each slice determines how many values contribute that slice's weight to the total sum. Sign handling then determines whether the contribution is positive or negative. Although this operation is not the main kernel target in the present transformer implementation, it demonstrates that the BSI representation supports a broader numeric algebra than just dot product.

\subsection{Element-Wise Multiplication}
Element-wise multiplication can also be expressed through bit-sliced interactions. The magnitude slices of one vector are combined with the magnitude slices of another vector, while the sign bitmap of the output is the exclusive-or of the two input sign bitmaps. This is more expensive than the corresponding dense operation, but it reinforces the key idea that the representation is capable of supporting arithmetic directly.

\section{Lessons from the Earlier CPU Evaluation}
The earlier CPU-side evaluation used bit-sliced vector operations on randomly generated vectors and on weights extracted from BERT \cite{bert,imdb}. The random-data experiments used vectors with up to ten million elements across multiple value ranges. The BERT experiments flattened trained model weights into one-dimensional arrays and compared BSI-based operations against dense PyTorch and C++ vector baselines. In both cases, the aim was to determine whether the representation could reduce storage cost and expose useful bit-level arithmetic structure before any GPU specialization was attempted.

Two findings from that earlier study shaped the present work directly.

First, BSI often reduced memory consumption, especially for dot product and summation. That made the representation attractive for resource-constrained inference settings. Second, CPU execution time remained poor for the core arithmetic operations, particularly for dot product and multiplication. The BSI method carried too much bit-manipulation and bookkeeping overhead on general-purpose CPU code to compete with optimized dense implementations. That outcome was the turning point of the project. It suggested that the representation itself remained promising, but that the arithmetic had to be moved to hardware more naturally suited to bit-parallel execution.

\section{Dense Quantization Context}
The present work is part of a broader inference-optimization landscape. Earlier work on limited-precision deep learning and accelerator co-design already showed that arithmetic width, model quality, memory traffic, and hardware mapping are tightly coupled \cite{gupta2015limitedprecision,lee2018unpu,mishra2020survey}. That line of work established the main design axes that still organize quantized LLM inference today: post-training quantization versus quantization-aware training, weight-only quantization versus joint weight-and-activation quantization, symmetric versus asymmetric scaling, per-tensor versus per-channel scaling, and static calibration versus input-dependent or dynamic rescaling. These choices determine not only numerical error but also whether the resulting tensors map cleanly onto dense accelerator kernels.

At LLM scale, one of the central numerical problems is the presence of outlier channels and heavy-tailed activation distributions. \texttt{LLM.int8()} addresses this by isolating outlier feature dimensions so that most of the matrix multiply can proceed in 8-bit arithmetic without allowing a small number of extreme channels to dominate the error budget \cite{llmint8}. ZeroQuant takes a different approach: it combines post-training quantization with layer-wise distillation and hardware-friendly integer execution, emphasizing an end-to-end deployment path rather than only an isolated quantizer \cite{zeroquant}. GPTQ frames the problem as an error-compensated post-training weight quantization problem and uses second-order information to preserve quality under aggressive low-bit weight compression \cite{gptq}.

AWQ and SmoothQuant push the same dense quantization theme in two complementary directions. AWQ uses activation statistics to identify salient channels and protects those channels from excessive quantization error, which allows the remaining weights to be quantized more aggressively without a corresponding collapse in model quality \cite{awq}. SmoothQuant instead observes that activation outliers are often the main obstacle to stable W8A8 inference and shifts part of that outlier burden from activations into weights through an offline smoothing transformation \cite{smoothquant}. This changes the distribution seen at inference time and makes dense weight-and-activation quantization more practical on contemporary GPU hardware.

These methods differ in calibration strategy and error model, but they share one structural property: the execution substrate remains dense. Quantization changes the values and scaling metadata, but the underlying kernel is still a dense matrix multiplication or convolution kernel tuned for integer or mixed-precision arithmetic. The problem studied here is different. The representation itself is bit-sliced, and the execution path is required to remain bit-sliced rather than converting the data back into a dense low-precision matrix before the kernel launch. The point of comparison is therefore not only accuracy under quantization, but whether a bit-sliced representation can support a practical inference pipeline when storage format, runtime builders, and kernel structure are all constrained to remain native to that representation.

\section{Transformer Context}
OPT \cite{opt} and the transformer architecture more broadly \cite{attention} provide a practical testbed for this study. Transformer inference stresses precisely the operations that BSI must accelerate: repeated dot products and matrix products in attention projections and feed-forward layers. This makes the transformer setting a natural place to ask whether the bit-sliced algebra that was useful on CPU can be scaled into a practical GPU inference pipeline.
